---
title: 'Template - Lab 3: Exploring Happiness'
subtitle: "Due: Thursday 10/03/2019 by 10 pm"
author: "Lynna Chu"

date: "09/26/2019"
output: html_document
---


## Background

The General Social Survey (GSS) has been run by NORC every other year since 1972 to keep track of current opinions across the United States. We are working with an excerpt of the dataset that covers years 1972 to 2016. The variable names along with a brief description are as follows: 

| Variable      | Description   |    
| ------------- |:-------------:| 
| Year          | Year for this respondent | 
| Age           | Age of respondent | 
| Sex           | Sex of respondent     | 
|....           | .....      | 

Now it's important to understand what each column stands for. You can find a codebook with explanations for each of the variables at https://gssdataexplorer.norc.org/

## Data Cleaning 

Initial exploration revealed that some data processing was needed before we could proceed with further visualization. All missing values were encoded with 'NA' as follows:
```{r}
happy <- readRDS("HAPPY.rds")
names(happy) = tolower(names(happy))
head(happy)
happy[happy == "NA"] = NA
happy[happy == "IAP"] = NA
happy[happy == "DK"] = NA
```

In total there were `....` number of missing values for the following `....` variables. 

Variables were also cast into the correct type in order to ease data visualization. This was done by:
```{r}
happy$age[happy$age == "89 OR OLDER"] = "89"
happy$age = as.numeric(happy$age)
factor(happy$happy)
factor(happy$sex)
factor(happy$marital)
factor(happy$degree)
factor(happy$finrela)
factor(happy$health)
factor(happy$wtssall)
factor(happy$partyid)
factor(happy$polviews)
```

Factors were also organized into meaningful order as follows: 
```{r}
#insert relevant R code for Q3 here
```


## Data Exploration 

Two variables of interest are happiness and age. Their distributions are as follow: 
```{r}
library(ggplot2)
ggplot(happy, aes(happy, fill = sex)) + geom_bar() + labs(title = "Barchart of Happiness", x = "Happy", y = "Count")
ggplot(happy, aes(age)) + geom_histogram(bins = 20) + labs(title = "Histogram of Age", x = "Age", y = "Count")
```

The distribution for happiness has a lot more people who are pretty happy or very happy compared to not too happy, suggesting that most people are fairly happy with their lives. There is an approximately even distribution of sex within each level of happiness, suggesting neither sex is more happy than the other.

The distibution for age is a skew right. Suggesting that more people are of the age 20-40 than those in the years older than that.




